---
project: "Utah_Child_Care"
script purpose: Script calculates UT 2025 child care costs
author: "Brian Son, Sarah Brolliar"
date: "2024-12-11"
validator: SBr on "2025-01-16" 
last modified: "2025-01-17"
last editor: "Sbr"
---

##Instructions:##
- Update file paths and variables as needed by state
- Ensure required libraries are installed

## Editing history
# 01/17/2025 - Add reminders for validators
# 01/16/2025 - Edits made to filepaths for movement to the new directory made by SBr. Add tests. Standardize output
# 01/16/2025 - Correct an error in the 2-3 year age cost file. 
# 01/28/2025 - LMa renamed columns in final output for alignment with County data file.

## Load Necessary Packages, Define File Paths
##### Change State info
```{r setup, include=FALSE}

state <- "utah"
state_abbrev <- "UT"
state_abbrev_lowercase <- "ut"
SSS_year <- "2025"

# Install necessary packages if not already installed
packages <- c("sssUtils", "tidyverse")
installed <- rownames(installed.packages())
missing_packages <- packages[!packages %in% installed]
if (length(missing_packages) > 0) {
  install.packages(missing_packages)
}

#Load libraries
library(sssUtils)  # Custom package
library(tidyverse) 
library(readxl)
library(tidyr)
library(dplyr)

#Set base paths
# Set base path
base_path <- get_base_path()
if (!dir.exists(base_path)) {
  stop("Base path does not exist. Please check your configuration.")
}
print(paste("Base path is:", base_path))

# Remove trailing slash from base_path if present
base_path <- sub("/$", "", base_path)

reference_path <- file.path(
  base_path, 
  "sss_production",
  "data",
  "reference")

base_path_raw <- file.path(
  base_path, 
  "sss_production",
  "data",
  "raw",
  "2025", 
  "child_care", 
  state)

base_path_processed <- file.path(
  base_path, 
  "sss_production",
  "data",
  "processed",
  "2025", 
  "child_care", 
  state)

output_path <- file.path(
  base_path, 
  "sss_production",
  "data",
  "processed",
  "2025", 
  "child_care")
```

## Notes for further adjustment

Define base CPI, Define raw data reported rate (daily, monthly, annual)

## State specific data

```{r notes}
#Correct landing page
CCDF_landing_page <- "https://jobs.utah.gov/occ/occmarket.pdf"

#raw_data
raw_data <- file.path(base_path_raw, "2024 Utah child care market rate study and narrow cost analysis.pdf")
metro_nonmetro_0_23_months <- file.path(base_path_processed, "0_23_metro_non-metro rates.xlsx") #Validated SBr, Table 3.7
metro_nonmetro_2_3_years <- file.path(base_path_processed, "2_3_metro_non-metro rates.xlsx") #Fixed, Validated, SBr, Table 3.8, 3.9
metro_nonmetro_4_5_years <- file.path(base_path_processed, "4_5_metro_non-metro rates.xlsx") #Validated SBr, Tables 3.10, 3.11
metro_nonmetro_6_12_years <- file.path(base_path_processed, "6_12_metro_non-metro rates.xlsx") #Validated SBr, Tables 3.12

state_geographies <- file.path(reference_path, "state_geo_list", "UT_geographies.csv")
region_county <- file.path(base_path_processed, "geo_county.xlsx")
region_county <- read_excel(region_county)
region_checked_by <- "SBr"
###########################################################Validator, check region to county

# The Designation of counties in regions was pieced together via 2021 MRS, and 2023 Cost Estimation Model
nonmetropolitan_region_source <- file.path(base_path_processed, "2021occmarket_source of metro areas.pdf")
#Section 5.2: The regions used for analysis in this section are defined by the Office of Management and Budget (OMB) and integrated into the statistical area definitions of the Bureau of Economic Analysis (BEA) (United States Bureau of Economic Analysis, n.d.). They include five metropolitan   statistical areas (MSAs) and one nonmetropolitan area. The MSAs are Logan (Cache County), Ogden-Clearfield (including Box Elder, Davis, Morgan and Weber Counties), Provo-Orem (including Juab and Utah Counties), Salt Lake (Salt Lake and Tooele Counties), and St. George (Washington County). The remaining counties comprise the nonmetropolitan portion of the state.

nonmetropolitan_region_source <- file.path(base_path_processed, "2023costmodel_source of nonmetro areas.pdf")
#Page 25: Figure 12 shows the monthly per-child cost of childcare by child age and CCQS Rating for the
#Logan, Ogden-Clearfield, Provo-Orem, Salt Lake City, and St. George metropolitan statistical
#areas, and the Central Nonmetropolitan Region (including Beaver, Garfield, Iron, Kane, Millard,
#Piute, Sanpete, Sevier, and Wayne Counties) and Eastern Nonmetropolitan Region (including
#Carbon, Daggett, Duchesne, Emery, Grand, Rich, San Juan, Summit, Uintah, and Wasatch
#Counties). The results are presented in ascending order of cost, with Logan at the left as the
#region with the lowest cost of care, and the Eastern Nonmetropolitan Region on the right with
#the highest cost of care.

#Survey inputs
#citation <- Ruetschlin, C. (2024). Utah 2024 Child Care Market Rate Study and Narrow Cost Analysis: Final Report prepared for the Utah Department of Workforce Services Office of Child Care. Salt Lake City, UT: University of Utah, Economic Evaluation Unit.
accessed_date <- 12/8/2024
accessed_by <- "Bso"
calculations_by <- "BsoSbr"
checked_by <- "Sbr"

survey_start_month <- "January"
survey_start_yr <- "2024"
survey_end_month <- "March"
survey_end_year <- "2024"
base_cpi_month <- "February"
base_cpi_year <- "2024"
cpi_checker <- "SBr"
###########################################################Validator, check cpi

#SSS Age Definitios
# infant = 0-2 years (up to 36 months)
# preschooler = 3-5 years
# school-age = 6-12 years

#Utah Age Definition Ratios
infant_0_23_months_ratio <- (2/3)
infant_2_year_old_ratio <- (1/3)
preschooler_3_year_old_ratio <- (1/3)
preschooler_4_year_old_ratio <- (1/3)
preschooler_5_year_old_ratio <- (1/3)
school_age_60_and_more_months <- 1
age_checker <- "SBr"
###########################################################Validator, check state ages

#All rates are reported as monthly, full-time do not need a monthly conversion 
#weekly_monthly <- 4.33
#hourly_monthly <- ___*5*(52/12)
#daily_monthly <- 5*(52/12)
monthly_conversion <- 1
rates_checker <- "SBr"
###########################################################Validator, check reported rates
```

## SSS Definitions, do NOT change per state

```{r SSS Definitions}
#infant
infant_center_weight <- .57
infant_home_weight <- .43

#preschooler
preschooler_center_weight <- .74
preschooler_home_weight <- .26

#school-age
school_center_weight <- .54
school_home_weight <- .46

#summer weight
avg_number_insruction_days <- 180
avg_number_work_days <- 262

#Percentage of workdays in school (before/after school care assumed)
before_after_school_care <- .69

#Percentage of summer/vacation school days (full time care assumed)
full_time_care <- .31

#Converting a fulltime school-age rate to a parttime school-age rate
parttime_conversion <- (5/9)
```

## Cleanup the dataset

```{r cleanup}
# Function to process and add suffix to columns and standardize 'Region' column
process_sheet_with_suffix <- function(df, age_group_suffix) {
  # Standardize the 'Region' column name
  df <- df %>%
    rename_with(~ ifelse(. %in% c("Region", "Regions"), "Region", .))
  
  # Check and rename 'Center Licensed' column
  if (any(grepl("^Center Licensed", colnames(df)))) {
    df <- df %>%
      rename_with(
        ~ paste0(., "_", age_group_suffix),
        starts_with("Center Licensed")
      )
  }
  
  # Check and rename 'Family Licensed' column
  if (any(grepl("^Family Licensed", colnames(df)))) {
    df <- df %>%
      rename_with(
        ~ paste0(., "_", age_group_suffix),
        starts_with("Family Licensed")
      )
  }
  
  return(df)
}

# Load datasets and process each with their respective suffix
metro_nonmetro_0_23_months <- read_excel(metro_nonmetro_0_23_months, sheet = 1) %>%
  process_sheet_with_suffix("0_23_months")

metro_nonmetro_2_years <- read_excel(metro_nonmetro_2_3_years, sheet = 1) %>%
  process_sheet_with_suffix("2_years")

metro_nonmetro_3_years <- read_excel(metro_nonmetro_2_3_years, sheet = 2) %>%
  process_sheet_with_suffix("3_years")

metro_nonmetro_4_5_years_sheet1 <- read_excel(metro_nonmetro_4_5_years, sheet = 1) %>%
  process_sheet_with_suffix("4_years")

metro_nonmetro_5_years <- read_excel(metro_nonmetro_4_5_years, sheet = 2) %>%
  process_sheet_with_suffix("5_years")

metro_nonmetro_6_12_years_sheet1 <- read_excel(metro_nonmetro_6_12_years, sheet = 1) %>%
  process_sheet_with_suffix("6_12_years")

# Combine datasets iteratively using a loop
sheet_list <- list(
  metro_nonmetro_0_23_months,
  metro_nonmetro_2_years,
  metro_nonmetro_3_years,
  metro_nonmetro_4_5_years_sheet1,
  metro_nonmetro_5_years,
  metro_nonmetro_6_12_years_sheet1
)

# Initialize the combined dataset with the first sheet
combined_data <- sheet_list[[1]]

# Loop through the remaining sheets and join by 'Region'
for (i in 2:length(sheet_list)) {
  combined_data <- full_join(combined_data, sheet_list[[i]], by = "Region")
}
```

## Calculate the data by age groups

```{r calculate the averages and Weighted SSS categories}
# Process combined data to calculate weighted averages and rename columns
combined_data <- combined_data %>%
  # Combine Infant Center
  mutate(
    Infant_Center = `Center Licensed_0_23_months` * infant_0_23_months_ratio +
      `Center Licensed_2_years` * infant_2_year_old_ratio,
    Infant_Family = `Family Licensed_0_23_months` * infant_0_23_months_ratio +
      `Family Licensed_2_years` * infant_2_year_old_ratio
  ) %>%
  select(-`Center Licensed_0_23_months`, -`Center Licensed_2_years`,
         -`Family Licensed_0_23_months`, -`Family Licensed_2_years`) %>%
  # Combine Preschooler Center
  mutate(
    Preschooler_Center = rowMeans(select(., 
                                         `Center Licensed_3_years`, 
                                         `Center Licensed_4_years`, 
                                         `Center Licensed_5_years`), 
                                  na.rm = TRUE),
    Preschooler_Family = rowMeans(select(., 
                                         `Family Licensed_3_years`, 
                                         `Family Licensed_4_years`, 
                                         `Family Licensed_5_years`), 
                                  na.rm = TRUE)
  ) %>%
  select(-`Center Licensed_3_years`, -`Center Licensed_4_years`, -`Center Licensed_5_years`,
         -`Family Licensed_3_years`, -`Family Licensed_4_years`, -`Family Licensed_5_years`) %>%
  # Rename School Age columns
  rename(
    `School_Age_Center_Fulltime` = `Center Licensed_6_12_years`,
    `School_Age_Family_Fulltime` = `Family Licensed_6_12_years`) 
```

##SSS Create school-age part time values from full time values, the school-age yearly values

```{r calculate the school-age parttime values}
 combined_data <- combined_data %>%
  # Define School Age parttime care
  mutate(
    `School_Age_Center_Parttime` = `School_Age_Center_Fulltime` * parttime_conversion, 
    `School_Age_Family_Parttime` = `School_Age_Family_Fulltime` * parttime_conversion) %>%
  #Define School-age monthly care
  mutate(
    `School_Age_Center_Monthly` = `School_Age_Center_Parttime` * before_after_school_care + 
      `School_Age_Center_Fulltime` * full_time_care, 
    `School_Age_Family_Monthly` = `School_Age_Family_Parttime`* before_after_school_care + 
      `School_Age_Family_Fulltime` * full_time_care) %>%
  #Drop extra schoolage columns
  select(-`School_Age_Family_Fulltime`, -`School_Age_Family_Parttime`, -`School_Age_Center_Fulltime`, -`School_Age_Center_Parttime`)
```

##Assign SSS Definitions ratios for Family and Center Licensed Calculation

```{r calculate the averages and Weighted SSS categories}
# Combine datasets iteratively using a loop
# (Previous steps for loading and combining datasets remain unchanged)

# Calculate SSS_School_Age
SSS_combined_data <- combined_data %>%
  # Step 1: Multiply `School_Age_Center` by school_center_weight
  mutate(School_Age_Center_Weighted = `School_Age_Center_Monthly` * school_center_weight) %>%
  
  # Step 2: Multiply `School_Age_Family` by school_home_weight
  mutate(School_Age_Family_Weighted = `School_Age_Family_Monthly` * school_home_weight) %>%
  
  # Step 3: Sum the weighted columns to create `SSS_School_Age`
  mutate(School_Age = School_Age_Center_Weighted + School_Age_Family_Weighted) %>%
  
  # Step 4: Remove intermediate and original columns
  select(-`School_Age_Center_Monthly`, -`School_Age_Family_Monthly`, 
         -School_Age_Center_Weighted, -School_Age_Family_Weighted)

# Calculate SSS_Preschooler
SSS_combined_data <- SSS_combined_data %>%
  # Step 1: Multiply `Preschooler_Center` by preschooler_center_weight
  mutate(Preschooler_Center_Weighted = `Preschooler_Center` * preschooler_center_weight) %>%
  
  # Step 2: Multiply `Preschooler_Family` by preschooler_home_weight
  mutate(Preschooler_Family_Weighted = `Preschooler_Family` * preschooler_home_weight) %>%
  
  # Step 3: Sum the weighted columns to create `SSS_Preschooler`
  mutate(Preschool = Preschooler_Center_Weighted + Preschooler_Family_Weighted) %>%
  
  # Step 4: Remove intermediate and original columns
  select(-`Preschooler_Center`, -`Preschooler_Family`, 
         -Preschooler_Center_Weighted, -Preschooler_Family_Weighted)

# Calculate SSS_Infant
SSS_combined_data <- SSS_combined_data %>%
  # Step 1: Multiply `Infant_Center` by infant_center_weight
  mutate(Infant_Center_Weighted = `Infant_Center` * infant_center_weight) %>%
  
  # Step 2: Multiply `Infant_Family` by infant_home_weight
  mutate(Infant_Family_Weighted = `Infant_Family` * infant_home_weight) %>%
  
  # Step 3: Sum the weighted columns to create `SSS_Infant`
  mutate(Infant = Infant_Center_Weighted + Infant_Family_Weighted) %>%
  
  # Step 4: Remove intermediate and original columns
  select(-`Infant_Center`, -`Infant_Family`, 
         -Infant_Center_Weighted, -Infant_Family_Weighted)

# Rename non-metro regions to align with joined dataframe. 
SSS_combined_data <- SSS_combined_data %>%
  mutate(
    Region = case_when(
      Region == "Central Nonmetropolitan" ~ "Central Nonmetropolitan Region",
      Region == "Eastern Nonmetropolitan" ~ "East Nonmetropolitan Region",
      TRUE ~ Region  # Keep other rows unchanged
    )
  )

```

## Assign costs to counties, and county fips codes.

```{r Join SSS weighted dataframe into region to county dataset}
# Load new dataset `regioncounty`
#region_county <- read_excel(region_county) #Already done above to allow for early validation

# Join new dataset onto `SSS_combined_data`
UT25_Childcare <- SSS_combined_data %>%
  left_join(region_county, by = "Region") %>%
  relocate(County, .before = Infant) %>%
  select(-Region)

state_geographies <- read_csv(state_geographies)

# Step 1: Rename the "County" column to "countyname" in `UT25_Childcare`
UT25_Childcare <- UT25_Childcare %>%
  rename(countyname = County)

# Step 2: Perform the join with `geo_county` using the renamed column
UT25_Childcare <- state_geographies %>%
  left_join(UT25_Childcare, by = "countyname")

# Step 3: Reorder columns so `Infant`, `Preschool`, and `School_Age` come right after `county_town_name`
final_result <- UT25_Childcare %>%
  relocate(Infant, Preschool, School_Age, .after = pop2022)

# Step 4: View the final output to verify
print(head(final_result))
```
# Update column names

```{r rename columns}

# Rename columns in the dataframe
final_result <- final_result %>%
  rename(
    infant = Infant,
    preschooler = Preschool,
    school_age = `School_Age`
  )


```
#Test for missing values

```{r test data}
# Check if 'infant' column has NA
any(is.na(final_result$infant))

# Check if 'preschooler' column has NA
any(is.na(final_result$preschooler))

# Check if 'school_age' column has NA
any(is.na(final_result$school_age))

```

#Save file

```{r}

# Get today's date in YYYY-MM-DD format
today_date <- Sys.Date()

# Define initials (replace "ABC" with your initials)
initials <- calculations_by

# Create the dynamic filename
filename <- paste0("processed_child_care_", state_abbrev_lowercase, "_", SSS_year, ".csv")

# Define the full file path
full_file_path <- file.path(output_path, filename)

# Save the final_region_county_list dataframe as a CSV file
write.csv(final_result, file = full_file_path, row.names = FALSE)

# Print confirmation
message("File saved as: ", filename)

```