---
project: "Texas_Child_Care"
script purpose: Script calculates TX 2025 child care costs
author: "Sarah Brolliar"
date created: "2024-11-04"
last modified: "2025-01-10"
last editor: "Sbr"
---

## 1.Setup-----------------

```{r setup, include=FALSE}

state <- "texas"
state_abbrev <- "TX"
state_abbrev_lowercase <- "tx"
SSS_year <- "2025"

# Install necessary packages if not already installed
packages <- c("sssUtils", "tidyverse")
installed <- rownames(installed.packages())
missing_packages <- packages[!packages %in% installed]
if (length(missing_packages) > 0) {
  install.packages(missing_packages)
}

# Load libraries
library(sssUtils)  # Custom package
library(tidyverse) 
library(readxl)
library(tidyr)
library(dplyr)
library(readr)

# Set base path
base_path <- get_base_path()
if (!dir.exists(base_path)) {
  stop("Base path does not exist. Please check your configuration.")
}
print(paste("Base path is:", base_path))

# Remove trailing slash from base_path if present
base_path <- sub("/$", "", base_path)

reference_path <- file.path(
  base_path, 
  "sss_production",
  "data",
  "reference")

base_path_raw <- file.path(
  base_path, 
  "sss_production",
  "data",
  "raw",
  "2025", 
  "child_care", 
  state)

base_path_processed <- file.path(
  base_path, 
  "sss_production",
  "data",
  "processed",
  "2025", 
  "child_care", 
  state)

output_path <- file.path(
  base_path, 
  "sss_production",
  "data",
  "processed",
  "2025", 
  "child_care")
 
#dir.create(output_path, showWarnings = FALSE)  # Create output directory if not present


```

## Editing history
```{r}
#1/10/2025 - Edits made to filepaths for movement to the new directory made by SBr. 
# 1/17/2025 = Validation by AKu
## moved the remove NA values from result up before the county concat in order to successfully remove the state of TX values, which were not being removed previously.
```
## State specific data
```{r notes}
#CCDF landing page = <https://txicfw.socialwork.utexas.edu/2024-texas-child-care-market-rate-survey/>

raw_data_source <- file.path(base_path_raw, "2024 Texas Child Care Market Rate Survey.pdf")
raw_data <- file.path(base_path_processed, "SupplementalReport_2024_Market_Rate_14Jun2024_nocalc_Rinput.xlsx")
state_geographies <- file.path(reference_path, "state_geo_list", "TX_geographies.csv")
county_assignments <- file.path(base_path_processed, "TX_LWDA_County.xlsx")
checked_by <- ""

#Survey inputs
citation <- "Schroeder, D., Marra, L. & Faulkner, M. (2024). 2024 Texas Child Care Market Rate Survey: Final Report prepared for the Texas Workforce Commission. Austin, TX: The University of Texas at Austin, Ray Marshall Center for the Study of Human Resources, Texas Institute for Child & Family Wellbeing"
accessed_date <- 10/31/2024
accessed_by <- "Sbr"
calculations_by <- "Sbr"
checked_by <- "Aku"

#Survey period
survey_start_month <- "September"
survey_start_yr <- "2023"
survey_end_month <- "May"
survey_end_year <- "2024"
Base_CPI_month <- "January"
Base_CPI_year <- "2024"
CPI_checker <- "Aku"

#state age definitions
infant_0_11_months_ratio <- (1/3)
infant_12_17_months_ratio <- (1/6)
infant_18_23_months_ratio <- (1/6)
infant_2_years_ratio <- (1/3)
preschooler_3years_ratio <- (1/3)
preschooler_4years_ratio <- (1/3)
preschooler_5years_ratio <- (1/3)
school_age_ratio <- 1

#TX has daily rates
reported_rate <- "daily"
#convert daily rates to monthly
monthly_conversion <- 5*(52/12)
#weekly_monthly <- 4.33
#hourly_monthly <- ___*5*(52/12)
#daily_monthly <- 5*(52/12)
```

## SSS Definitions, do NOT change per state

```{r SSS Definitions}
#infant
infant_center_weight <- .57
infant_home_weight <- .43

#preschooler
preschooler_center_weight <- .74
preschooler_home_weight <- .26

#school-age
school_center_weight <- .54
school_home_weight <- .46

#summer weight
avg_number_insruction_days <- 180
avg_number_work_days <- 262

#Percentage of workdays in school (before/after school care assumed)
before_after_school_care <- .69

#Percentage of summer/vacation school days (full time care assumed)
full_time_care <- .31

#Converting a fulltime school-age rate to a parttime school-age rate
parttime_conversion <- (5/9)
```

## Cleanup the dataset

```{r cleanup}

# Load the Excel file
df <- read_excel(raw_data) 

# Rename columns for clarity (optional but helpful)
colnames(df) <- c("Location", "Facility_Type", "Age_Group", "Age_Years", "Care_Type", "Mean", "30th_Percentile", 
                  "40th_Percentile", "Median", "60th_Percentile", "70th_Percentile", "75th_Percentile")

# Remove rows where Location is NA
df_2 <- df %>%
  filter(!is.na(Location))

#Drop unneeded columns
df_3 <- df_2 %>%
  select(-"Age_Group", -"Mean", -"30th_Percentile", -"40th_Percentile", -"Median", -"60th_Percentile", -"70th_Percentile")

# Fill missing Age_Group values downwards
df_4 <- df_3 %>% fill(Facility_Type, Age_Years, .direction = "down")
```

## Widen the dataset, trimming extra columns as you go

```{r widen}
# Convert to wide format
  #Care_type, or part time vs. full time care, is the variable brought out of the column into the row
df_wide_care_type <- df_4 %>%
  pivot_wider(
    names_from = Care_Type,
    values_from = c("75th_Percentile")
  )

#Shorten the child care facility labels for ease
df_wide_care_type <- df_wide_care_type %>%
  mutate(Facility_Type = substr(Facility_Type, 1, 25))

df_wide_care_type <- df_wide_care_type %>%
  mutate(Facility_Type = case_when(
    Facility_Type == "Licensed Child Care Cente" ~ "Center",
    Facility_Type == "Registered Child Care Hom" ~ "R_Home",
    Facility_Type == "Licensed Child Care Homes" ~ "L_Home",
    TRUE ~ Facility_Type  # Keeps other values unchanged
  ))

# Convert to wide format
  #Bring the age_years out of the column into the row
df_wide_age_years <- df_wide_care_type %>%
  pivot_wider(
    names_from = "Age_Years",
    values_from = c("Full-day", "Part-day")
  )

#Drop all part-time care columns for anyone not a school child
df_wide_age_years <- df_wide_age_years %>%
  select(-'Part-day_0-11 months', -'Part-day_12-17 months', -'Part-day_18-23 months', -'Part-day_2 years', - 'Part-day_3 years', -'Part-day_4 years', -'Part-day_5 years') 

```

## Convert data into SSS age categories

```{r convert to SSS ages}
#Weight ages to fit into SSS age classifications: 
  #infant: 1/3*0-11 months, 1/6*12-17 months, 1/6*18-23 months, 1/3*2years
  #preschooler: 1/3*3 years, 1/3*4 years, 1/3*5 years
  #school-age: 6-13 years

#First, ensure that they are numeric
df_SSS_ages <- df_wide_age_years %>%
  mutate(
    `Full-day_0-11 months` = as.numeric(`Full-day_0-11 months`),
    `Full-day_12-17 months` = as.numeric(`Full-day_12-17 months`),
    `Full-day_18-23 months` = as.numeric(`Full-day_18-23 months`),
    `Full-day_2 years` = as.numeric(`Full-day_2 years`),
    `Full-day_3 years` = as.numeric(`Full-day_3 years`),
    `Full-day_4 years` = as.numeric(`Full-day_4 years`),
    `Full-day_5 years` = as.numeric(`Full-day_5 years`),
    `Full-day_6-13 years` = as.numeric(`Full-day_6-13 years`),
    `Part-day_6-13 years` = as.numeric(`Part-day_6-13 years`),

# Now create the SSS age categories
infant_full_daily = infant_0_11_months_ratio * `Full-day_0-11 months` + infant_12_17_months_ratio * `Full-day_12-17 months` + infant_18_23_months_ratio * `Full-day_18-23 months` + infant_2_years_ratio * `Full-day_2 years`, 

preschooler_full_daily = preschooler_3years_ratio * `Full-day_3 years` + preschooler_4years_ratio * `Full-day_4 years` +  preschooler_5years_ratio * `Full-day_5 years`,

schoolage_full_daily = `Full-day_6-13 years`,
schoolage_part_daily = `Part-day_6-13 years`
)

#Drop month and year specific columns, leaving only SSS age categories
df_SSS_ages <- df_SSS_ages %>%
  select(-`Full-day_0-11 months`, -`Full-day_12-17 months`, -`Full-day_18-23 months`, 
         -`Full-day_2 years`, -`Full-day_3 years`, -`Full-day_4 years`, 
         -`Full-day_5 years`, -`Full-day_6-13 years`, -`Part-day_6-13 years`)
```

## Combine schoolage part and full time care (weighted)

```{r combine schoolage part and full}
# Weight part and full time care for school-age children into a single value
    # Confirm columns are numeric
    # Weight the full and part time care for schoolage children
    # Remove the unneeded full time and part time columns 
df_SSS_ages_weight_schoolage_care <- df_SSS_ages %>%
  mutate(
    schoolage_full_daily = as.numeric(schoolage_full_daily),
    schoolage_full_daily = as.numeric(schoolage_part_daily),
    schoolage_combinedcare_daily = (full_time_care * schoolage_full_daily) + (before_after_school_care * schoolage_part_daily)
  )

df_SSS_ages_weight_schoolage_care <- df_SSS_ages_weight_schoolage_care %>%
  select(- schoolage_full_daily, - schoolage_part_daily)
```

## Convert daily to monthly costs

```{r daily to monthly}
# Convert daily costs into Monthly costs
  #remove the daily cost columns
df_SSS_ages_monthly <- df_SSS_ages_weight_schoolage_care %>%
  mutate(
    schoolage_combinedcare_monthly = schoolage_combinedcare_daily*monthly_conversion,
    infant_full_monthly = infant_full_daily*monthly_conversion,
    preschooler_full_monthly = preschooler_full_daily*monthly_conversion 
  )
df_SSS_ages_monthly <- df_SSS_ages_monthly %>%
select(- 'infant_full_daily', - 'preschooler_full_daily', -'schoolage_combinedcare_daily')
```

## Widen the data set to get weighted child care costs by SSS age and facility ratios, trimming as you go

```{r widen for SSS age and facility}
# Convert to wide format
# Bring the facility type out of the column and into one row
df_SSS_ages_monthly_wide_facility_type <- df_SSS_ages_monthly %>%
  pivot_wider(
    names_from = "Facility_Type",
    values_from = c("schoolage_combinedcare_monthly", "infant_full_monthly", "preschooler_full_monthly")
  )

# Combine the two home costs by SSS age group
# Texas has registered and licensed homes. Average them together, unweighted.
df_SSS_ages_monthly_center_home <- df_SSS_ages_monthly_wide_facility_type %>%
  mutate(
    infant_full_monthly_home = (infant_full_monthly_L_Home + infant_full_monthly_R_Home) / 2,  
    preschooler_full_monthly_home = (preschooler_full_monthly_L_Home + preschooler_full_monthly_R_Home) / 2,
    schoolage_combinedcare_monthly_home = (schoolage_combinedcare_monthly_R_Home + schoolage_combinedcare_monthly_L_Home) / 2
  )

# Drop the separated home columns
df_SSS_ages_monthly_center_home <- df_SSS_ages_monthly_center_home %>%
  select(-infant_full_monthly_L_Home, 
         -infant_full_monthly_R_Home, 
         -preschooler_full_monthly_L_Home, 
         -preschooler_full_monthly_R_Home, 
         -schoolage_combinedcare_monthly_R_Home, 
         -schoolage_combinedcare_monthly_L_Home)

# Weight SSS age costs by facility
df_SSS_childcare_monthly_weightedfacility <- df_SSS_ages_monthly_center_home %>%
  mutate(
    infant = infant_full_monthly_Center * infant_center_weight + infant_full_monthly_home * infant_home_weight, 
    preschooler = preschooler_full_monthly_Center * preschooler_center_weight + preschooler_full_monthly_home * preschooler_home_weight, 
    school_age = schoolage_combinedcare_monthly_Center * school_center_weight + schoolage_combinedcare_monthly_home * school_home_weight
  )

# Drop the separate facility columns
df_SSS_childcare_monthly_weightedfacility <- df_SSS_childcare_monthly_weightedfacility %>%
  select(-infant_full_monthly_Center, 
         -infant_full_monthly_home, 
         -preschooler_full_monthly_Center, 
         -preschooler_full_monthly_home, 
         -schoolage_combinedcare_monthly_Center, 
         -schoolage_combinedcare_monthly_home)
```

#Join the datasets
```{r}
# Convert Location to County
location_to_county <- read_excel(county_assignments, sheet = 1)

# Rename the LWDA column to Location
df_SSS_childcare_monthly_weightedfacility <- df_SSS_childcare_monthly_weightedfacility %>%
  rename(LWDA = Location)

# Join the dataframes
result <- df_SSS_childcare_monthly_weightedfacility %>%
  left_join(location_to_county, by = "LWDA")

# Remove the LWDA column
result <- result %>%
  select(-LWDA)

# Reorder columns to make County the first column
result <- result %>%
  relocate(County, .before = everything())

# Remove rows with NA values
TX2025 <- result %>%
  drop_na()

# Add " County" to each entry in the County column and rename it to "countyname"
TX2025 <- TX2025 %>%
  mutate(County = paste0(County, " County")) %>%
  rename(countyname = County)



# View the result
head(result)

```

## Assign regions to their counties, and county fips codes.
# Test for NA values. 

```{r assign regions to counties}

# Read the data from region_to_county
state_geographies <- read_csv(state_geographies)
state_geographies <- as.data.frame(state_geographies)

# Match SSS data costs to the region and county key
final_result <- state_geographies %>%
  left_join(TX2025, by = "countyname")

# Check if 'infant' column has NA
any(is.na(final_result$infant))

# Check if 'preschooler' column has NA
any(is.na(final_result$preschooler))

# Check if 'school_age' column has NA
any(is.na(final_result$school_age))

```

```{r}
# Get today's date in YYYY-MM-DD format
today_date <- Sys.Date()

# Define initials (replace "ABC" with your initials)
initials <- calculations_by

# Create the dynamic filename
filename <- paste0("processed_child_care_", state_abbrev_lowercase, "_", SSS_year, ".csv")

# Define the full file path
full_file_path <- file.path(output_path, filename)

# Save the final_region_county_list dataframe as a CSV file
write.csv(final_result, file = full_file_path, row.names = FALSE)

# Print confirmation
message("File saved as: ", filename)
```