---
title: "Georgia_Child_Care"
author: "Sarah Brolliar"
date: "2024-11-04"
output: html_document
---

## Editing history
# 2/10/2025 - Convert script to new format with base paths, added tests, naming conventions, etc. Renamed variables without spaces. Made school-age part time calculation easier to identify. Checked values. 

## Load Necessary Packages

```{r setup, include=FALSE}
state <- "georgia"
state_abbrev <- "GA"
state_abbrev_lowercase <- "ga"
SSS_year <- "2025"

# Install necessary packages if not already installed
packages <- c("sssUtils", "tidyverse")
installed <- rownames(installed.packages())
missing_packages <- packages[!packages %in% installed]
if (length(missing_packages) > 0) {
  install.packages(missing_packages)
}

# Load libraries
library(sssUtils)  # Custom package
library(tidyverse) 
library(readxl)
library(tidyr)
library(dplyr)
library(readr)

# Set base path
base_path <- get_base_path()
if (!dir.exists(base_path)) {
  stop("Base path does not exist. Please check your configuration.")
}
print(paste("Base path is:", base_path))

# Remove trailing slash from base_path if present
base_path <- sub("/$", "", base_path)

reference_path <- file.path(
  base_path, 
  "sss_production",
  "data",
  "reference")

base_path_raw <- file.path(
  base_path, 
  "sss_production",
  "data",
  "raw",
  SSS_year, 
  "child_care", 
  state)

base_path_processed <- file.path(
  base_path, 
  "sss_production",
  "data",
  "processed",
  SSS_year, 
  "child_care", 
  state)

output_path <- file.path(
  base_path, 
  "sss_production",
  "data",
  "processed",
  SSS_year, 
  "child_care")
 
#dir.create(output_path, showWarnings = FALSE)  # Create output directory if not present

```

## Notes for further adjustment

Define base CPI Define raw data reported rate (daily, monthly, annual)

## State specific data

```{r notes}
#Correct landing page
#CCDF landing page = https://www.decal.ga.gov/bfts/ResearchMarketRates.aspx

#raw_data
table_1 <- file.path(base_path_processed, "Table 1.xlsx")
table_2 <- file.path(base_path_processed, "Table 2.xlsx")
table_3 <- file.path(base_path_processed, "Table 3.xlsx")
table_4 <- file.path(base_path_processed, "Table 4.xlsx")
table_5 <- file.path(base_path_processed, "Table 5.xlsx")
table_6 <- file.path(base_path_processed, "Table 6.xlsx")
table_key <- file.path(base_path_processed, "Table_key.xlsx")

#state_path <- 'G:/Shared Drives/CWW Team Drive/SSS/All States/GA/GA 2025/'
state_geographies <- file.path(reference_path, "state_geo_list", "GA_geographies.csv")

#There are no indications of changes to county assignments from the 2021 survey to the 2024 survey. 2021 survey included a map, and discussion of counties whose designations have changed. The absence of similar sections in the 2024 supports the same zone to county assignments for GA. 
region_to_county <- file.path(base_path_processed,"GA_region_county.xlsx")
region_to_county <- read_excel(region_to_county)
region_checked_by <- ""

#Survey inputs
#citation <- "Schroeder, D., Marra, L. & Faulkner, M. (2024). Georgia Child Care Market Rate Data 2023
accessed_date <- 10/22/2024
accessed_by <- "Sbr"
calculations_by <- "Sbr"
checked_by <- "Bso"

#"For this analysis, the Georgia Department of Early Care and Learning (DECAL) provided contractor Care Solutions, Inc. Excel spreadsheets with provider information as of October 2023 with a total of 4,425 child care learning center and family child care learning home records, 3,435 of which (78%) had rates that were confirmed or updated in 2023." 
survey_start_month <- " "
survey_start_yr <- " "
survey_end_month <- "October"
survey_end_year <- "2023"
print(survey_CPI_month <- "October")
print(survey_CPI_year <- "2023")

#state age definitions
##0-11 months = infant = .33
##12 - 17 months = infant = .166667
##18 - 23 months = infant = .166667
##2 years = infant = .33
## 3 years = preschooler = .33
## 4 years = preschooler = .33
## 5 years preschooler = .33
##6 - 13 years = school-age = 1.0

#Missing Data: No homes reported rates for 5 year olds. Preschool was taken as an average of 3 and 4 years olds for home sites, with the edits made manually in the code below.

infant_0_11_months <- .33
infant_12_23_months <- .33
infant_24_35_months <- .33
preschooler_3years <- .33
preschooler_4years <- .33
preschooler_5years <- .33
school_age <- 1

#All rates are reported as weekly, full-time. 
#convert weekly to monthly
#daily_monthly <- 5*(52/12)
#weekly_monthly <- 4.33
monthly_conversion <- 4.33
```

## SSS Definitions, do NOT change per state

```{r SSS Definitions}
#infant
infant_center_weight <- .57
infant_home_weight <- .43

#preschooler
preschooler_center_weight <- .74
preschooler_home_weight <- .26

#school-age
school_center_weight <- .54
school_home_weight <- .46

#summer weight
avg_number_insruction_days <- 180
avg_number_work_days <- 262

#Percentage of workdays in school (before/after school care assumed)
before_after_school_care <- .69

#Percentage of summer/vacation school days (full time care assumed)
full_time_care <- .31

#Converting a fulltime school-age rate to a parttime school-age rate
parttime_conversion <- (5/9)
```

## Cleanup the dataset

```{r cleanup}

# Load the Excel file
table_1 <- read_excel(table_1) 
table_2 <- read_excel(table_2) 
table_3 <- read_excel(table_3) 
table_4 <- read_excel(table_4) 
table_5 <- read_excel(table_5) 
table_6 <- read_excel(table_6) 
table_key <- read_excel(table_key)

# Rename columns for clarity (optional but helpful)
# Later - pursue code bringing the entries for the 1st row entries up as the column names
# Later - pursue code to rename the columns of all 6 tables as once
colnames(table_1) <- c("rate_type", "n", "mean", "std_deviation", "minimum", "maximum", "25th_Percentile", 
                  "50th_Percentile", "75th_Percentile")
colnames(table_2) <- c("rate_type", "n", "mean", "std_deviation", "minimum", "maximum", "25th_Percentile", 
                  "50th_Percentile", "75th_Percentile")
colnames(table_3) <- c("rate_type", "n", "mean", "std_deviation", "minimum", "maximum", "25th_Percentile", 
                  "50th_Percentile", "75th_Percentile")
colnames(table_4) <- c("rate_type", "n", "mean", "std_deviation", "minimum", "maximum", "25th_Percentile", 
                  "50th_Percentile", "75th_Percentile")
colnames(table_5) <- c("rate_type", "n", "mean", "std_deviation", "minimum", "maximum", "25th_Percentile", 
                  "50th_Percentile", "75th_Percentile")
colnames(table_6) <- c("rate_type", "n", "mean", "std_deviation", "minimum", "maximum", "25th_Percentile", 
                  "50th_Percentile", "75th_Percentile")

table_1a <- table_1 %>%
select(-"n", -"mean", -"std_deviation", -"minimum", -"maximum", -"25th_Percentile", -"50th_Percentile")
table_2a <- table_2 %>%
select(-"n", -"mean", -"std_deviation", -"minimum", -"maximum", -"25th_Percentile", -"50th_Percentile")
table_3a <- table_3 %>%
select(-"n", -"mean", -"std_deviation", -"minimum", -"maximum", -"25th_Percentile", -"50th_Percentile")
table_4a <- table_4 %>%
select(-"n", -"mean", -"std_deviation", -"minimum", -"maximum", -"25th_Percentile", -"50th_Percentile")
table_5a <- table_5 %>%
select(-"n", -"mean", -"std_deviation", -"minimum", -"maximum", -"25th_Percentile", -"50th_Percentile")
table_6a <- table_6 %>%
select(-"n", -"mean", -"std_deviation", -"minimum", -"maximum", -"25th_Percentile", -"50th_Percentile") 

#Drop CAPS rows and quality rating rows from all tables
table_1b <- table_1a[-(11:23),]
table_2b <- table_2a[-(11:23),]
table_3b <- table_3a[-(11:23),]
table_4b <- table_4a[-(11:23),]
table_5b <- table_5a[-(11:23),]
table_6b <- table_6a[-(11:23),]
#Drop first 3 rows with header info from all tables
table_1c <- table_1b[-(1:3),]
table_2c <- table_2b[-(1:3),]
table_3c <- table_3b[-(1:3),]
table_4c <- table_4b[-(1:3),]
table_5c <- table_5b[-(1:3),]
table_6c <- table_6b[-(1:3),]

```

#Confirm transformations, delete extra dataframes

```{r confirm transformation}

table_1 <- table_1c
table_2 <- table_2c
table_3 <- table_3c
table_4 <- table_4c
table_5 <- table_5c
table_6 <- table_6c

rm(table_1a, table_1b, table_1c, table_2a, table_2b, table_2c, table_3a, table_3b, table_3c, table_4a, table_4b, table_4c, table_5a, table_5b, table_5c, table_6a, table_6b, table_6c)

```
# Convert to wide format
```{r}

### Since the tables are separated nicely and neatly, changes are being made to indicate facility type and zone using a key. 
# However, since I am operating without internet and ChatGPT to solve dumb syntax issues, for now I will enter it semi-manually. 
# Later - insert code to use the table_key to populate the correct facility and zone
# Add column indicating facility. Populate tables 1-3 with "center", 4-6 "home" 


# Convert to wide format, bring the facility out of the column into the row
for (table_num in 1:6) {
  # Construct the table name
  table_name <- paste0("table_", table_num)
  
  # Check if the table exists in the global environment
  if (exists(table_name)) {
    # Get the table
    df <- get(table_name)
    
    # Convert to wide format
    df <- df %>%
      pivot_wider(
        names_from = "rate_type",
        values_from = c("75th_Percentile")
      )
    
    # Save the modified table back to the global environment
    assign(table_name, df)
    
    # Print confirmation
    message(paste("Converted", table_name, "to wide format."))
  } else {
    message(paste("Table", table_name, "does not exist in the global environment."))
  }
}
```


#Write a function to rename variables in a clean way
```{r function clean}

clean_column_names <- function(df) {
  df %>%
    rename_with(~ gsub("\\s+$", "", .x) %>%    # Remove trailing spaces
                  gsub(" ", "_", .) %>%        # Replace spaces with _
                  tolower() %>%                # Convert to lowercase
                  gsub("-", "_", .))          # Convert - to _
}

table_1 <- clean_column_names(table_1)
table_2 <- clean_column_names(table_2)
table_3 <- clean_column_names(table_3)
table_4 <- clean_column_names(table_4)
table_5 <- clean_column_names(table_5)
table_6 <- clean_column_names(table_6)



```

## Create school-age part time values
```{r}


parttime_values <- function(df) {
  df %>%
    mutate(
    school_age_rate = as.numeric(`school_age_rate`), # Convert to numeric if needed
    school_part_rate = school_age_rate * parttime_conversion
  )
}

table_1 <- parttime_values(table_1)
table_2 <- parttime_values(table_2)
table_3 <- parttime_values(table_3)
table_4 <- parttime_values(table_4)
table_5 <- parttime_values(table_5)
table_6 <- parttime_values(table_6)

```


## Convert data into SSS age categories - center only. Home done in next chunk to account for missing data for 5 year olds. Combine schoolage part and full time care (weighted)

```{r convert to SSS ages}
#Weight ages to fit into SSS age classifications: 
# Loop through Tables 1 to 3 - Center 
for (table_num in 1:3) {
  # Construct the table name as a string, e.g., "Table_22"
  table_name <- paste0("table_", table_num)
  
  # Check if the table exists in the global environment
  if (exists(table_name)) {
    # Retrieve the table from the global environment
    df <- get(table_name)
    
    # Ensure all relevant columns are numeric
    df[] <- lapply(df, function(x) as.numeric(as.character(x)))
    

# Now create the SSS age categories, by weighting rates by proportion of age group
df <- df %>%
      mutate(
        infant = (infant_0_11_months * `under_1_year_rate`) +
                      (infant_12_23_months * `1_year_old_rate`) +
                      (infant_24_35_months * `2_year_old_rate`),
        preschooler = (preschooler_3years * `3_year_old_rate`) +
                           (preschooler_4years * `4_year_old_rate`) +
                           (preschooler_5years * `5_year_old_rate`),
        schoolage_full = school_age * `school_age_rate`,
        schoolage_part = school_age * `school_part_rate`,
        school_age = (before_after_school_care * schoolage_part) + (full_time_care * schoolage_full)
      )


#Drop month and year specific columns, leaving only SSS age categories
df <- df %>%
  select(-`under_1_year_rate`, -`1_year_old_rate`, -`2_year_old_rate`, 
         -`3_year_old_rate`, -`4_year_old_rate`, -`5_year_old_rate`, -`school_age_rate`, -`schoolage_part`, -`schoolage_full`, -`school_part_rate`)

 # Save the modified table back to the global environment
    assign(table_name, df)
    
    # Print confirmation
    message(paste("Table:", table_name, "| Calculations added and columns dropped"))
  } else {
    message(paste("Table", table_name, "does not exist in the global environment"))
  }
}
```

##Same as above for home tables. Combine schoolage part and full time care (weighted)
```{r convert to SSS ages}
#Weight ages to fit into SSS age classifications: 
# Loop through Tables 1 to 6
for (table_num in 4:6) {
  # Construct the table name as a string, e.g., "Table_22"
  table_name <- paste0("table_", table_num)
  
  # Check if the table exists in the global environment
  if (exists(table_name)) {
    # Retrieve the table from the global environment
    df <- get(table_name)
    
    # Ensure all relevant columns are numeric
    df[] <- lapply(df, function(x) as.numeric(as.character(x)))
    

# Now create the SSS age categories
df <- df %>%
      mutate(
        infant = (infant_0_11_months * `under_1_year_rate`) +
                      (infant_12_23_months * `1_year_old_rate`) +
                      (infant_24_35_months * `2_year_old_rate`),
        preschooler = (.5 * `3_year_old_rate`) + (.5 * `4_year_old_rate`),
        schoolage_full = school_age * `school_age_rate`,
        schoolage_part = school_age * `school_part_rate`,
        school_age = (before_after_school_care * schoolage_part) + (full_time_care * schoolage_full)
      )


#Drop month and year specific columns, leaving only SSS age categories
df <- df %>%
  select(-`under_1_year_rate`, -`1_year_old_rate`, -`2_year_old_rate`, 
         -`3_year_old_rate`, -`4_year_old_rate`, -`5_year_old_rate`, -`school_age_rate`, -`schoolage_part`, -`schoolage_full`, -`school_part_rate`)

 # Save the modified table back to the global environment
    assign(table_name, df)
    
    # Print confirmation
    message(paste("Table:", table_name, "| Calculations added and columns dropped"))
  } else {
    message(paste("Table", table_name, "does not exist in the global environment"))
  }
}
```

#Add suffixes to label home vs. center

```{r }
#Add suffixes from the table of contents to each Table. Instead of 75th percentile, if the ToC tables says Table 22 is by Cost Area 
#and center, then I can add cost area and center to 75th percentile to differentiate. This is done in two steps

#Add facility suffix
# Loop through Tables 1 to 6
for (table_num in 1:6) {
  # Construct the table name as a string, e.g., "Table_22"
  table_name <- paste0("table_", table_num)
  
  # Filter `ToC` for the current table number and retrieve the suffix from 'age_group'
  suffix <- table_key %>%
    filter(table_number == table_num) %>%  # Use the correct column name `table_number`
    slice(1) %>%                          # Select only the first matching row
    pull(facility_type)                       # Pull the value from 'age_group'
  
  # Check if we found exactly one unique suffix
  if (length(suffix) == 1 && !is.na(suffix)) {
    # Check if the table exists in the global environment
    if (exists(table_name)) {
      # Get the table from the global environment
      df <- get(table_name)
      
       # Create new column names for all columns
      new_col_names <- paste0(colnames(df), "_", suffix)
      
      # Rename all columns
      colnames(df) <- new_col_names
      
      # Save the modified table back to the global environment
      assign(table_name, df)
      
      # Print to check that it's working
      message(paste("Table:", table_name, "| Suffix found:", suffix))
    } else {
      message(paste("Table", table_name, "does not exist in the global environment"))
    }
  } else {
    message(paste("No unique suffix found in ToC for Table", table_name))
  }
}

```

#Join center and home tables by region. Calculate the SSS ages of center and home separately first before joining by region to do the final facility weights. 

```{r}

#Add column indicating zone. Tables 1, 4 are zone 1, 2, 5 are zone 2, and 3 and 6 are zone 3. 
table_1$region <- 1
table_4$region <- 1
table_2$region <- 2
table_5$region <- 2
table_3$region <- 3
table_6$region <- 3

#Join tables by region
# Make a more concise, automatic code using the table_key for this later. 
region_1 <- table_1 %>%
  left_join(table_4, by = "region")
region_2 <- table_2 %>%
  left_join(table_5, by = "region")
region_3 <- table_3 %>%
  left_join(table_6, by = "region")


#First, ensure that they are numeric
region_3 <- region_3 %>%
  mutate(across(everything(), ~ as.numeric(as.character(.))))

# Check the structure to confirm
str(region_3)

```
# Apply facility weights 
```{r}
#Facility ratios

# Define regions (assumes regions are named as df_region_1, df_region_2, etc.)
region_numbers <- 1:3

# Loop through each region
for (region_num in region_numbers) {
  # Construct the region dataframe name as a string, e.g., "df_region_1"
  region_name <- paste0("region_", region_num)
  
  # Check if the regional dataframe exists in the global environment
  if (exists(region_name)) {
    # Retrieve the dataframe from the global environment
    df <- get(region_name)
    
    # Ensure all relevant columns are numeric
    df[] <- lapply(df, function(x) if(is.character(x) || is.factor(x)) as.numeric(as.character(x)) else x)
    
      # Apply calculations
      df <- df %>%
        mutate(
          infant = (infant_center_weight * infant_Center) +
                   (infant_home_weight * infant_Home),
          preschooler = (preschooler_center_weight * preschooler_Center) +
                        (preschooler_home_weight * preschooler_Home),
          school_age = (school_center_weight * school_age_Center) +
                       (school_home_weight * school_age_Home)
        )
      
    #Drop facility specific columns, leaving only SSS age categories
    df <- df %>%
  select(-`infant_Center`, -`preschooler_Center`, -`school_age_Center`, 
         -`infant_Home`, -`preschooler_Home`, -`school_age_Home`)
    
      # Save the modified dataframe back to the global environment
      assign(region_name, df)
      
      # Print confirmation
      message(paste("Region:", region_name, "| Costs weighted by facility ratios applied"))
    } else {
    # If the region dataframe does not exist
    message(paste("Region", region_name, "does not exist in the global environment"))
  }
}
```
## Convert to monthly costs

```{r monthly conversion}
#Facility ratios

# Define regions (assumes regions are named as df_region_1, df_region_2, etc.)
region_numbers <- 1:3

# Loop through each region
for (region_num in region_numbers) {
  # Construct the region dataframe name as a string, e.g., "df_region_1"
  region_name <- paste0("region_", region_num)
  
  # Check if the regional dataframe exists in the global environment
  if (exists(region_name)) {
    # Retrieve the dataframe from the global environment
    df <- get(region_name)
    
    # Ensure all relevant columns are numeric
    df[] <- lapply(df, function(x) if(is.character(x) || is.factor(x)) as.numeric(as.character(x)) else x)
    
#Apply calculations
    df <- df %>%
      mutate(
        infant = (monthly_conversion * infant),
        preschooler = (monthly_conversion * preschooler), 
        school_age = (monthly_conversion * school_age)
      )
    
      # Save the modified dataframe back to the global environment
      assign(region_name, df)
      
      # Print confirmation
      message(paste("Region:", region_name, "| Costs weighted by facility ratios applied"))
    } else {
    # If the region dataframe does not exist
    message(paste("Region", region_name, "does not exist in the global environment"))
  }
}
```

## Combine into one dataframe

```{r combine}
# Combine all region dataframes into one
combined_regions <- bind_rows(region_1, region_2, region_3, .id = "region_id")

# View the combined dataframe
print(combined_regions)

```

## Assign regions to their counties, and county fips codes.

```{r assign regions to counties}

GA25_county_costs <- region_to_county %>%
  mutate(region = as.numeric(region)) %>% # Convert region to numeric
  left_join(combined_regions, by = "region")

#Drop the extra columns
GA25_county_costs <- GA25_county_costs %>%
  select(-"Zone", -"region", -"region_id")

# Generate the final input of county with FIPS and costs 
# Bring in the county
county_list <- read_csv(state_geographies)

final_result <- county_list %>%
  left_join(GA25_county_costs, by = c("countyname" = "County"))
```


#Test for missing values
```{r test data}
# Check if 'infant' column has NA
any(is.na(final_result$infant))

# Check if 'preschooler' column has NA
any(is.na(final_result$preschooler))

# Check if 'school_age' column has NA
any(is.na(final_result$school_age))
```

```{r}

# Get today's date in YYYY-MM-DD format
today_date <- Sys.Date()

# Define initials (replace "ABC" with your initials)
initials <- calculations_by

# Create the dynamic filename
filename <- paste0("processed_child_care_", state_abbrev_lowercase, "_", SSS_year, ".csv")

# Define the full file path
full_file_path <- file.path(output_path, filename)

# Save the final_region_county_list dataframe as a CSV file
write.csv(final_result, file = full_file_path, row.names = FALSE)

# Print confirmation
message("File saved as: ", filename)


```